---
title: "Temporal Records"
author: "Ben Weinstein"
date: "Thursday, January 08, 2015"
output:
  html_document:
    theme: spacelab
---

#Aim

One of the fundamental questions associated with climate change is how plants will respond to rising temperatures. To accurately model plant phenology, we need to consider both the error associated 

#Approach


```{r}
library(dplyr)
library(reshape2)
library(ggplot2)
```

#What's wrong with linear regression

##simulate data

Consider a simple time series model where phenology is advancing due to rising temperatures. The advancment is modeled as a poisson process where lambda is the julian date from the previous year.

$$Y_i \sim y_{i-1} - N(\mu=1,sd=50)$$


```{r, echo=FALSE}
years<-seq(1901,2000)

#trajectory
d<-function(){
days<-c()
days[1]<-180
for (x in 2:length(years-1)){
  days[x]<- days[x-1] - rnorm(n=1,mean=1,sd=50)
}
out<-data.frame(days,years)
return(out)
}

dat<-rbind_all(replicate(1000,d(),simplify=F))

q<-group_by(dat,years) %>% summarize(mean=mean(days))

p<-ggplot(q,aes(years,mean)) + geom_point() + theme_bw() + ylab("First Flowering Date")
p
```

#Fit a linear line

If we assume that the first flowering date is normally distributed, then we can model the data as:

$$Y_i \sim N(\mu,\sigma)$$
$$\mu = \alpha + \beta * year$$

```{r}
m<-lm(data=q,mean~years)
summary(m)
p+ geom_line(aes(y=predict(m,newdata=q)),col="Red")
```
